---
title: "Washington State Cancer Analysis"
author: "David Chiang, Nancy Liu, Roy Mosby, Jeff Tu"
date: "5/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../top_cancers.R")
library(plotly)
```
## Introduction
In this current pandemic time in the start of 2020 where many people are falling
victim to COVID-19, we want to raise awareness while educating ourselves on
the potential health burdens that other diseases can bring to people. Specifically,
we will narrow down our scope to people affected by cancer in Washington state.
We will examine how the mortality rates by cancer differ by cause, county,
gender, and more within the state of Washington.


## Relationship Between Cancer and Mortality Rate 
```{r age, echo = FALSE, message = FALSE}
ggplotly(top_10_cancers, tooltip = c("text"))
```
In order to measure the serverity of different types of cancer, we can visualize
the mortality rate of the top 10 cancers using a bar chart. Here, we analyze
the relationship between the independent variable of type of cancer and the
dependent variable of mortality rate. The mortality rate measures the number of
deaths in a particular population, and is expressed here in units of death per
1,000 individuals per year. The origin of this data source, the Institute for
Health Metrics and Evaluation (IHME) as referenced in our reference list
at the end of this page, collected this data from 1980 - 2014. This bar chart
represents the average of the recorded mortality rate of each case from this
time period. We see that the highest mortality rate is close to 50 for 
tracheal, bronchus, and lung cancer. The IHME has categorized these three types
of cancer into one likely because the symptoms and effects are similar to one
another. This group leads the mortality rate by cancer by a wide margin with
most types of cancer having a mortality rate of between 5 and 20. The
bar chart being left skewed suggests that most cancers do not have a mortality
rate that drastically differs from one another apart from the lung cancer group.

## References
http://ghdx.healthdata.org/record/ihme-data/united-states-cancer-mortality-rates-county-1980-2014?fbclid=IwAR1_4xWKPLDf1Pl5JzD40zDEScrIwCvPGTCmeRRlwtbYDYjR5v1aZIIOLdo
