---
title: "Washington State Cancer Analysis"
author: "David Chiang, Nancy Liu, Roy Mosby, Jeff Tu"
date: "5/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("../top_cancers.R")
source("../mean_mortality.R")
source("../ethnic_rates.R")
library(plotly)
```
## Introduction
In this current pandemic time in the start of 2020 where many people are falling
victim to COVID-19, we want to raise awareness while educating ourselves on
the potential health burdens that other diseases can bring to people. Specifically,
we will narrow down our scope to people affected by cancer in Washington state.
We will examine how the mortality rates by cancer differ by cause, county,
gender, and more within the state of Washington.


## Relationship Between Cancer and Mortality Rate 
```{r age, echo = FALSE, message = FALSE}
ggplotly(top_10_cancers, tooltip = c("text"))
```
In order to measure the severity of different types of cancer, we can visualize
the mortality rate of the top 10 cancers using a bar chart. Here, we analyze
the relationship between the independent variable of type of cancer and the
dependent variable of mortality rate. The mortality rate measures the number of
deaths in a particular population, and is expressed here in units of death per
1,000 individuals per year. The origin of this data source, the Institute for
Health Metrics and Evaluation (IHME) as referenced in our reference list
at the end of this page, collected this data from 1980 - 2014. This bar chart
represents the average of the recorded mortality rate of each case from this
time period. We see that the highest mortality rate is close to 50 for 
tracheal, bronchus, and lung cancer. The IHME has categorized these three types
of cancer into one likely because the symptoms and effects are similar to one
another. This group leads the mortality rate by cancer by a wide margin with
most types of cancer having a mortality rate of between 5 and 20. The
bar chart being left skewed suggests that most cancers do not have a mortality
rate that drastically differs from one another apart from the lung cancer group.

<<<<<<< HEAD
## Mean Mortality Rates of the Top Ten Cancers
=======
## Mean Mortality Rate Distribution of Top Ten Cancers
>>>>>>> f1ee1b97182656fcbe686babd83303c50759852f
```{r year, echo = FALSE, message = FALSE}
ggplotly(year_mortality)
```
We analyzed the mean mortality rates coming from the deadliest cancers. In this 
case, we compiled the means by using the data from 40 counties in Washington
state. From this, we can tell that the overall range of the mortality rates has 
decreased when comparing the years 1980 and 2014. This decrease in range could 
be attributed to the fact that new technological advances in imaging, detection,
and treatment has allowed for cancer to be detected and treated earlier, 
allowing for people to survive these cancers. We can see that mortality rates
are decreasing for the following cancers: tracheal, bronchus, and lung cancer,
colon and rectum cancer, breast cancer, and prostate cancer. These cancers are
the most commonly heard of ones, and because of interventions such as campaigns 
against smoking, a wider set of genetic tests, we can see a drastic decrease in 
the mortality rates. Other cancers such as pancreatic, leukemia, brain and 
nervous system, and esophageal remain steady. In some cases such as brain and 
nervous system cancer, if a tumor grows on an inoperable location, not much intervention 
could help the patient be cured of this cancer. Thus even with advancements, the
mortality rate in these cases does not change significantly. This suggests 
that while certain types of cancers are seeing a decrease in mortality rates,
others are seeing either steady mortality rates or increasing mortality rates.

<<<<<<< HEAD
## Counties with the Highest Mean Mortality Rates in WA
```{r year, echo = FALSE, message = FALSE}
ggplotly(counties_cancer, tooltip = c("text"))
```
In order to better understand how cancer affected different counties in WA, we
looked at the mortality rates for cancer among all the counties in WA from 1980 to
2014. We compiledthe mean mortality rates for every county in WA, and observed the 
10 counties with the highest mortality rates. From this we can observe that the 
county with the greatest mortality rate was Grays Harbor County, with a mortality 
rate of approximately 15.63%. Aside from Grays Harbor County, the mortality
rates for the other counties with the highest mortality rates 
=======
## Rates of Top 3 Cancers by Race/Ethnicity 
```{r, echo = FALSE, message = FALSE, warning=FALSE}
ggplotly(cr_plot, tooltip = c("text"))
```
With the existence of so many different types of cancers. We wanted to explore
if a determining factor in cancer incidence could be linked to race/ethnicity.
Therefore we decided to create a distribution graph that presents the top 3 rates
of cancers by race/ethnicity. With Washington state having uneven racial
demographics, we used a technique called age adjustment, which is used to show
the rates that would exist if the population had the same distribution. Our
results show a significant difference between different ethnic groups and we can
make an hypothesis that this could be the cause of hereditary genetics. The data
that was used to create this distribution was made available by the Centers for
Disease Control and Prevention (CDC).


## Counties with the Highest Mean Mortality Rates in WA
```{r year, echo = FALSE, message = FALSE}
ggplotly(counties_cancer, tooltip = c("text"))
```
In order to better understand how cancer affected different counties in WA, we
looked at the mortality rates for cancer among all the counties in WA from 1980 to
2014. We compiledthe mean mortality rates for every county in WA, and observed the 
10 counties with the highest mortality rates. From this we can observe that the 
county with the greatest mortality rate was Grays Harbor County, with a mortality 
rate of approximately 15.63%. Aside from Grays Harbor County, the mortality
rates for the other counties with the highest mortality rates were approximately
14-15%, and are not significantly different from each other. 

>>>>>>> f1ee1b97182656fcbe686babd83303c50759852f
## References
http://ghdx.healthdata.org/record/ihme-data/united-states-cancer-mortality-rates-county-1980-2014?fbclid=IwAR1_4xWKPLDf1Pl5JzD40zDEScrIwCvPGTCmeRRlwtbYDYjR5v1aZIIOLdo

https://gis.cdc.gov/Cancer/USCS/DataViz.html

